---
title: "R Notebook"
output: html_notebook
---

```{r}

if (!require("tidyr")) {
  install.packages("tidyr", repos="http://cran.rstudio.com/") 
  library(tidyr)
}

if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}

if (!require("readxl")) {
  install.packages("readxl", repos="http://cran.rstudio.com/") 
  library(readxl)
}

if (!require("stringr")) {
  install.packages("stringr", repos="http://cran.rstudio.com/") 
  library(stringr)
}
```

```{r}
data3<-read_xlsx("C:/R-Source/data_dre.xlsx", sheet="Hoja1")
```



I.EXTRACT CHARACTERES FROM A COLUMN
==========================
#1. Get the parenthesis and what is inside
```{r}
ara.1$newSis_Con<-str_extract(ara.1$Sistema_consulta,"\\([^()]+\\)")
ara.1$newSis_Rep<-str_extract(ara.1$Sistema_rep01,"\\([^()]+\\)")
ara.1$newSis_Ana<-str_extract(ara.1$Sistema_Analisis,"\\([^()]+\\)")
```
 
 
#2. Remove parenthesis and rescue those who have not parenthesis 
```{r}
ara.1$newSis_Con <- substring(ara.1$newSis_Con,2, nchar(ara.1$newSis_Con)-1)
ara.1$newSis_Con <-ifelse(is.na(ara.1$newSis_Con),ara.1$Sistema_consulta,ara.1$newSis_Con)
```
 


II.REMOVE CHARACTERES INSIDE A COLUMN
===================================
#1.Remove point decimal
RT11: Remove  point decimal en campos de esfuerzo RTC
```{r RT11: Convirtiendo el punto decimal en campos de esfuerzo}
rtc.1 <- rtc.1%>%  
  separate(Esfuerzo_SFS_1,c("Esfuerzo_SFS_01","dec_sfs1"), sep ='\\.', remove = TRUE)

```


#2.remove accents
```{r}
ara.1$Area<-chartr("?????", "aeiou",ara.1$Area)  
```
 


III.UPDATE CHARACTERS

#1.uPDATE a whole dataframe column to 
```{r}
data<-data3[,c("Tiempo del Caso","Tiempo de  Solución Real","Tiempo de  Atención Real")]

data[] <- lapply(data, function(x) as.numeric(as.character(x)))
```



IV.DEDUPLICATION
====================================
#1.Separate characters united by ".",";"(point, semicolon,etc )
Using "Separate" from "tidyr library" with Dataframe
separate the colum "JerarquIa" into two columns splitted "(", it is necessary put "\\" before adding more conditions using "|".
```{r RT02:Separate "Jerarquia" from ARANDA}

data3.sep <- data3%>%  
  separate(Jerarquia,c("Macroservicio","Servicio_ti","Subservicio","Categoria1","Categoria2","Categoria3","Categoria4"), sep ='\\.|\\.|\\.|\\.|\\.|\\.', remove = FALSE)

``` 

#2.SPLIT "unid_org" into two column without deleting the original one
```{r}
ara.1 <- ara.1%>%  
  separate(Unid_org,c("Area","SubArea"), sep ='\\/',remove = FALSE)
```


 
 