---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
Install Libraries
```{r include=FALSE}
install.packages("readr")
install.packages("tidyverse")
install.packages("tidyr")
install.packages("dplyr")
install.packages("stringr")
install.packages("lubridate")

```


Calling Libraries
```{r Calling Libraries}
#library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)

#working with dates
library(lubridate)
```



IMPORTING
```{r}
#We can read from all "sheet" of any excel file
data1<-read_xlsx("C:/R-Source/procesos1.xlsx", sheet = "Hoja1")

```



I.SELECT ROW AND COLUMNS (dplyr PACKAGE)
====================================
##1.Basic Filter, Select and Arrange ("dplyr")
```{r}  
#Select basic
 select(iris, Petal.Length, Species)

#Select with pipe
iris %>%
          select(Petal.Length, Species)
        
iris %>%
          select(Petal.Length, Species)%>%
          filter(!Petal.Length%in%c(1.4,1.5))

#Arrange: order rows and assign to a object called "examSelec1"
examSelec1<-iris %>%
          select(Petal.Length, Species)%>%
          filter(Petal.Length%in%c(1.4,1.5))%>%
           arrange(desc(Petal.Length))
```

##2.Multiple condition
```{r}
df %>% filter(A == 1 & B == 3 | A == 3 & B ==2)
```


##3. Advance filters to identify into a word using "str_detect" ("stringr" library)
```{r}
#Using "fruit" dataset from "stringr" library
stringr::fruit
fruta=as.data.frame(fruit)
colnames(fruta)[1]="nombre"

##1. Finding the words that start with "black" or "blu"
fruta %>%
  filter(str_detect(nombre, 'black|blu'))

##2. Finding the words that start with "l"
fruta %>%
  filter(str_detect(nombre, "^l"))

##3. Finding the words that ends with "l"
fruta %>%
  filter(str_detect(nombre, "l$"))


##4.pick up rows where the names contain digits
  df_wrangling <- data.frame(name=c("Melody2","Gian","Paul44","Jearson","Jenny"),scores=c(3,1,5,6,3),Menbership=c(11,2,6,7,8))

df_wrangling %>% 
  filter(str_detect(name, "\\d"))

##5.pick up rows where the names NO contain digits
df_wrangling %>% 
  filter(!str_detect(name, "\\d"))

```

##4.Advance filters to identify words into a phrase using "grepl" ("base" library)
```{r}
library(dplyr)
stringr::sentences

frases=as.data.frame(sentences)
colnames(frases)[1]="name"


frases %>% filter(grepl("The so", name))
```



III.OPERATIONS
==========================
### 1.COUNT
1.Group_by and Summarize
```{r}
#without filter
iris %>%
    group_by(Petal.Length) %>%
    summarize(cantidad=n())

#without filter
iris %>%
    filter(Petal.Length%in%c(1.4,1.7))%>%
    group_by(Petal.Length) %>%
    summarize(cantidad=n())

```

### 2.SUM
```{r}
df.saea.clean%>%
          filter(terminado=="si"&Estado=="En Proceso"&grepl("INDRA",Especialista.de.Estado)&!is.na(serv_TI))%>%           group_by(Linea.de.Servicio=serv_TI,Fecha.final)%>%
          summarise(horas=round((sum(Tiempo)/60),2))

mtcars %>%
    mutate(NA_count = rowSums(is.na(.)))


df %>%
   summarise(across(everything(), ~ sum(., is.na(.), 0)))
```



##2. Tally
```{r}
# tally() is short-hand for summarise()
mtcars %>% tally()
# count() is a short-hand for group_by() + tally()
mtcars %>% count(cyl)

# add_tally() is short-hand for mutate() adding the number of records
mtcars %>% add_tally()

# add_count() is a short-hand for group_by() + add_tally(), adding the partial sum of a different colum value
mtcars %>% add_count(cyl)

# count and tally are designed so that you can call
# them repeatedly, each time rolling up a level of detail
species <- starwars %>% count(species, homeworld, sort = TRUE)
species
species %>% count(species, sort = TRUE)

```


### 2. Mean
```{r}
df.saea.clean%>%
          filter(terminado=="si"&Estado=="En Proceso"&grepl("INDRA",Especialista.de.Estado)&!is.na(serv_TI))%>% 
          group_by(Linea.de.Servicio=serv_TI,Fecha.final)%>%
          summarise(horas=round((sum(Tiempo)/60),2))%>%
          group_by(Linea.de.Servicio) %>%
          summarise(Mean = mean(horas))
```


III.UPDATE COLUMNS(dplyr, stringr PACKAGE)
=========================
##1.Update colums itself
1.1 using Mutate(dplyr)  
```{r}
iris %>%
  mutate(Species= ifelse(grepl("setosa",Species),"valor de prueba",Species)) 
```


##2.from another, prior to Compare one numeric column 
```{r}
#working with df.iris, created before

#create anohter dataframe
newspecie<-data.frame(newname=c("tulip","rose"),number=c(1,2))

#Update "df.iris", prior to compare against "newspecie"
#1ra way:
df.iris$Species<-ifelse(match(newspecie$number,df.iris$id),newspecie$newname,NA)


#2da way
df.iris$Species<-newspecie$newname[match(df.iris$id,newspecie$number)]
```

##3.from another, prior to Compare one string column 
```{r}
set.seed(1)
df1 <- data.frame(A = sample(seq(3), 10, replace = TRUE))
df2 <- data.frame(A = seq(3), B = c("pizza", "hot dog", "hamburger"), stringsAsFactors = FALSE)

#it is better use a simple match
df1$B <- df2$B[match(df1$A, df2$A)]
```


##4.Update one table from another, prior to Compare 2 columns
```{r}
rtc.1$estado_req<-ifelse(match(rtc.1$macroservicio,rtc.est.gen$Macroservicio_aux)&match(rtc.1$Estado,rtc.est.gen$Estado_req_aux),rtc.est.gen$Estado_generico,NA)
```

##5.Update words
```{r}

#You use ifelse when you have to replace all 
ara.dqa.dep$area<- ifelse(str_detect(ara.dqa.dep$area,regex("Direccion de Inversiones")),"DIN",ara.dqa.dep$area)
#if you just want to update same pattern without affect the others one, do that
ara.dqa.dep$area<- ifelse(str_detect(ara.dqa.dep$area,regex("Direccion de Inversiones")),"DIN",ara.dqa.dep$area)
```


III.ADD COLUMNS (Mutate)
===================================
##1.Add column for all rows
```{r}
#
vector1 %>%
  mutate(newColumn=ifelse(grepl(3.5,Sepal.Width),10,Sepal.Width)) 
```

##2.Add a colum to a group
```{r}
if (!require("gcookbook")) {
  install.packages("gcookbook", repos="http://cran.rstudio.com/") 
  library(gcookbook)
}

library(ggplot2)
library(dplyr)

#Add a column that catch the percentage as "ratio"
res <- cabbage_exp %>%
group_by(Date) %>%
mutate(ratio = Weight / sum(Weight))

#plot 
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
geom_bar(stat = "identity",colour = "black", position = "fill") +
geom_text(data = res, aes(label=scales::percent(ratio), y=ratio), position = position_fill(vjust = 0.5))

```


IV.SEPARATE COLUMS (PARSING)
=======================

##1.Separate columns in R by space delimiter
```{r}
#Creating DATA FRAME
df<-data.frame(technology=c("SOA PowerBuilder Java","PLSQL Python OT"),issues=c(2,8))
#Removing Whitespace using: \\s+
demo<-df%>%
    separate(technology,c("tech1","tech2","tech3"),sep = "\\s+",remove = FALSE)
```


##2.Separate into multiple columns
```{r}
df_separate<-data.frame(service_name=c("Macroservice1.Service1.Subservice1.CategoryA1.CategoryA2.CategoryA3.CategoryA4","Macroservice5.Service4.Subservice9.CategoryA9.CategoryA12.CategoryA23.CategoryA8"),issues=c(2,8))

#Split with special character(*,/,(, etc)), then use "separate"
df_separate%>%  
  separate(service_name,c("Macroservicio","Servicio_ti","Subservicio","Categoria1","Categoria2","Categoria3","Categoria4"), sep ='\\.|\\.|\\.|\\.|\\.|\\.', remove = TRUE)
  
```




V.JOIN (Base,purr PACKAGE)
========================
##1.Inner join
```{r}
  
df.bdi<-inner_join(list2,df.p2,by="n_requerimiento")
```



##2.inner join based on two columns
```{r}
library(dplyr)
df1 <- data.frame(
  lat = c(-33.9174, -33.9175, -33.9176, -33.9177, -33.9171), 
  long = c(151.2263, 151.2264, 151.2265, 151.2266, -140.54),
  PM = c(8, 10, 9, 8, 55)
)

df2 <- data.frame(
  lat = c(-33.9174, -33.9175, -33.9176, -33.9177, -31), 
  long = c(151.2263, 151.2264, 151.2265, 151.2266, 134),
  PM = c(12, 15, 11, 3, 18)
)


inner_join(df1, df2, by = c("lat", "long"))
```



```{r}
#1.CREATE COLUM
#For listing in vertical view the values of a "String" or "dataframe", it is necesary use "fill = 1"
cat(letters,fill = 1)


```

