---
title: "R Notebook"
output: html_notebook
---
```{r}
install.packages("nycflights13")

```

```{r}
library(nycflights13)

library(stringr)
library(stringi)
library(dplyr)
```


1. EXTRACT characters (Pattern)
==============================
##1. Extracts before or after a character 

###1.before or after a hyphen
```{r}
#Extracts characters before a hyphen
v2<-str_extract('Lima-Peru', "[^-]+")

```



##2. Extracts after, before space
```{r}
library(stringr)
shopping_list <- c("service S4", "loan for school", "loan for university", "Service S2")
```


```{r}
#1. before space
str_extract(shopping_list, "[a-z]+")
##the same
gsub( "\\s.*", "", shopping_list )
##

#str_extract(shopping_list, "[a-z]")

#2.First word after space
str_extract(shopping_list, " [a-z]+")


##3. Between spaces
str_extract("d1 service s1", "(?<=\\s)(.*)(?=\\s)")

#(?<=\\s) checks whether there is whitespace before
#(?=\\s) checks whether there is a whitespace after
#(.*) extracts everything in between the white spaces
```




```{r}
#. You can find the complete list here, but here we can use '[:alpha:]' to represent all alphabet characters including both upper and lower cases.

flight %>%
  select(FL_DATE, CARRIER, ORIGIN, ORIGIN_CITY_NAME, ORIGIN_STATE_ABR, DEP_DELAY, DEP_TIME, ARR_DELAY, ARR_TIME) %>%
filter(str_detect(ORIGIN_CITY_NAME, "New[:alpha:]+rk")) %>%
count(ORIGIN_CITY_NAME)
```


##3. Extracts only numbers
```{r}
shopping_list <- c("service S4", "loan for school", "loan for university", "Service S2")
```

```{r}
str_extract(shopping_list, "\\d")
  
```

##4 Extracts characters base on position

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
```


###1.Extract the first to four
```{r}
 str_extract(shopping_list, "[a-z]{1,4}")
```

###2.Extract the last four library(stringi)
```{r}

library(stringi)
stri_sub(shopping_list, -4)

```

```{r}
esfu.ini$id_padre=substr(esfu.ini$id_padre,1,nchar(esfu.ini$id_padre)-3)
```


###3.extr characters that start with b in position:  from the first to four
```{r}

str_extract(shopping_list, "\\b[a-z]{1,4}\\b")
```

##5.Extract dates 
```{r}

#Extract dates with this pattern "18/09/20" 
# datos_ingreso=c("Recuento: 1 Ultimo: 17/12/19 10")

 str_extract(datos_ingreso, "[0-9]{2}/[0-9]{2}/[0-9]{2}")
# 


```

II.EXTRACT A PATTERN
=======================

```{r}
  #Now we are going to use a "sentences" from "stringr" library
  stringr::sentences
  length(sentences)
  head(sentences)
  #I flip the order of the second and third words
  sentences%>%
          str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2")%>%
          head(5)
```



III. REMOVE
============================
##1.Removing vowels
```{r}

var4 <- c("1551.55", "AA two tables", "a cloud")

#remove vowels  
str_remove(var4, "[aeiou]")

#all
str_remove_all(var4, "[aeiou]")

```
##2.removing letters
```{r}
gsub(' [A-z ]*', '' , shopping_list)
```


##3.Remove all character after the first comma

```{r}
pattern <- data.frame("colid" = c(1, 2), 
                      "text" = c("john smith, jane smith", "jon stewart, steven colbert"))

df3=pattern %>% 
  mutate(text2 = str_replace_all(text, ",.*", ""))
```

##4.Remove all character after the first space
```{r not proved}

df3 %>% 
    mutate(text3 = sub("\\s+[^ ]+$", "", text2))

##the same
df %>% 
   mutate(name = str_replace(name, "\\s[^ ]+$", ""))

```


##5.Removing first character 
```{r}
str_remove(var4, ".")
```

##6.Removing dot (.)
```{r}
gsub("\\.", "", var4, perl=TRUE)
```


