---
title: "R Notebook"
output: html_notebook
---

```{r}
if (!require("purrr")) {
  install.packages("purrr", repos="http://cran.rstudio.com/") 
  library(purrr)
}

if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}

if (!require("tibble")) {
  install.packages("tibble", repos="http://cran.rstudio.com/") 
  library(tibble)
}
```


Basic Loop with purr
======================
##Map as lapply()
lapply() is basically identical to map() , except that map() is consistent with all the other
functions in purrr, and you can use the shortcuts for .f 
```{r}
df <- tibble(
a = rnorm(10),
b = rnorm(10),
c = rnorm(10),
d = rnorm(10)
)
```

map() makes a list.
map_lgl() makes a logical vector.
map_int() makes an integer vector.
map_dbl() makes a double vector.
map_chr() makes a character vector.

```{r}
map_dbl(df, mean)
#the same
df %>% map_dbl(mean)
```
```{r}
z <- list(x = 1:3, y = 4:5)
map_int(z, length)
```


##Shortcuts
```{r}
models <- mtcars %>%
split(.$cyl) %>%
map(function(df) lm(mpg ~ wt, data = df))

#Is the same that
models <- mtcars %>%
split(.$cyl) %>%
map(~lm(mpg ~ wt, data = .))

```

## reduce
```{r}
dfs <- list(
age = tibble(name = "John", age = 30),
sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)
```

```{r}
vs <- list(
c(1, 3, 5, 6, 10),
c(1, 2, 3, 7, 8, 10),
c(1, 2, 3, 4, 8, 9, 10)
)
vs %>% reduce(intersect)
```


##ACUMULATE- PURRR library
```{r}
x <- sample(10)
#
x %>% accumulate(`+`)
```

https://emoriebeck.github.io/R-tutorials/purrr/

```{r}
1:3 %>% accumulate("+")
1:10 %>% accumulate_right("*")

#using .init
1:10 %>% accumulate(max, .init = 5) #.init: make sure that no value will be less than 5
```
# Understanding the arguments .x and .y when .f is a lambda function and x is the accumulating value
```{r}
1:10 %>% accumulate(~ .x)
# .y is element in the list
1:10 %>% accumulate(~ .y)
```

```{r}
# Simulating stochastic processes with drift
## Not run:
library(dplyr)
library(ggplot2)
rerun(5, rnorm(100)) %>%
set_names(paste0("sim", 1:5)) %>%
map(~ accumulate(., ~ .05 + .x + .y)) %>%
map_dfr(~ data_frame(value = .x, step = 1:100), .id = "simulation") %>%
ggplot(aes(x = step, y = value)) +
geom_line(aes(color = simulation)) +
ggtitle("Simulations of a random walk with drift")
```



```{r}
dat <- tibble(
         day = 200:210,
         x = sample(-10:10, size = 11,replace = T))

# accumulate, do the recursive sum of 'x' elements while initiating with a value of 5 (.init = 5) and remove the first element of accumulate output ([-1])
dat %>%
     mutate(y = accumulate(x, ~ .x + .y, .init = 5)[-1])

```




