---
title: "R Notebook"
output: html_notebook
---

```{r Data library}

if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}


if (!require("kableExtra")) {
  install.packages("kableExtra", repos="http://cran.rstudio.com/") 
  library(kableExtra)
}


if (!require("knitr")) {
  install.packages("knitr", repos="http://cran.rstudio.com/") 
  library(knitr)
}

```



I.Basis
==========
#1.Creating a Table Directly
```{r}
m.table <- matrix(c(51,43,22,92,28,21,68,22,9),ncol=3,byrow=TRUE)
colnames(m.table) <- c("High","Low","Middle")
rownames(m.table) <- c("current","former","never")
m.table <- as.table(m.table)
```


#2.Addition and subtract
```{r}
#https://stackoverflow.com/questions/4946873/how-do-i-add-a-row-to-a-data-frame-with-totals

#https://garthtarr.github.io/meatR/janitor.html

#1.addition
x <- data.frame(Language=c("C++", "Java", "Python"), 
                Files=c(4009, 210, 35), 
                LOC=c(15328,876, 200), 
                stringsAsFactors=FALSE) 

y <- rbind(x, c("Total", colSums(x[,2:3])))



#2.Subtract
df <- read.table(textConnection("Nrow  a   b    c   d
1     0.1 0.2 0.3 0.4 
2      2   3   4   5
3      7   8   9   10"), stringsAsFactors=FALSE, header=TRUE)

df[1, ] - df[2, ]
# or assign the result to a new row...
df <- rbind(df, df[1, ] - df[2, ])

```

#3.Subtract from the previous colum
```{r}
data = data.frame(x=c(1,5,7,10),y=c(3,2,4,1.1),z=1:4)
```
##1Basic
```{r}
 for(i in 2:nrow(data)){
 result = data[i,] - data[1,]
 print (result)
 }
```
##2.LIke Cash flow
https://rpubs.com/DanielSLee/IntroRforFinance
```{r}
 for(i in 2:ncol(data)){
 result1 = data[i,] - data[1,]
 print (result)
 }
```



II.OPERATIONS
=========
##1.transpose A DATAFRAME
```{r}
library(data.table)

# get data
  data("mtcars")

# transpose
  t_mtcars <- transpose(mtcars)

# get row and colnames in order
  colnames(t_mtcars) <- rownames(mtcars)
  rownames(t_mtcars) <- colnames(mtcars)
  
```

##2.converting multiple columns from character to numeric format in r
```{r}
DF <- data.frame("a" = as.character(0:5),
                 "b" = paste(0:5, ".1", sep = ""),
                 "c" = letters[1:6],
                 stringsAsFactors = FALSE)

# Check columns classes
sapply(DF, class)

#           a           b           c 
# "character" "character" "character" 

cols.num <- c("a","b")
DF[cols.num] <- sapply(DF[cols.num],as.numeric)
sapply(DF, class)

```

##3.converting all columns- applied to an object of class "c('integer', 'numeric')"
```{r}
library(dplyr)
mtcars %>%
  mutate_all(as.character)
```

##4.converting all columns to integer
```{r}
data <- structure(list(v1 = c(22.394, 43.72, 58.544, 56.877, 1.659, 29.142, 
67.836, 68.851), v2 = c(144.373, 72.3, 119.418, 112.429, 35.779, 
41.661, 166.941, 126.548), v3 = structure(c(33L, 29L, 33L, 5L, 
13L, 31L, 5L, 8L), .Label = c("", "#VALUE!", "0", "1", "10", 
"11", "12", "13", "14", "15", "16", "17", "18", "19", "2", "20", 
"21", "22", "23", "24", "25", "26", "28", "29", "3", "30", "32", 
"33", "4", "48", "5", "6", "7", "8", "9"), class = "factor"), 
    v4 = structure(c(24L, 6L, 22L, 23L, 16L, 22L, 23L, 26L), .Label = c("", 
    "-1", "-2", "-4", "#VALUE!", "0", "1", "10", "11", "12", 
    "13", "14", "15", "16", "17", "18", "19", "2", "24", "28", 
    "29", "3", "4", "5", "6", "7", "8", "9"), class = "factor")), .Names = c("v1", 
"v2", "v3", "v4"), row.names = c("4", "5", "6", "7", "8", "9", 
"10", "11"), class = "data.frame")

data[] <- lapply(data, function(x) as.numeric(as.character(x)))
```

##5.count NAs per row in dataframe
```{r}
#df is a dataframe with NA values
df$na_count <- apply(df, 1, function(x) sum(is.na(x)))
```


##6Charging colum names to a "list"
```{r}
#de.1 is a data frame
lista1<-names(de.1)
#or viceversa also run, as for pareto bean
```


III.Cumulative values in a table
================
```{r}
### create a data frame of totals
    product_sales <- data.frame("Day" = c(1:10),"Sales" = c(100,200,150,120,130,145,90,150,180,190))
 
### add cumulative column to the data frame
    product_sales[,"cum_sales"] <- cumsum(product_sales$Sales)
 
###Add mean    
    product_sales[,"mean"] <- mean(product_sales$Sales)
    
### print data frame
    product_sales
```



##2.Cumulative frequency
```{r}

if (!require("agricolae")) {
  install.packages("agricolae", repos="http://cran.rstudio.com/") 
  library(agricolae)
}


weight<-c( 68, 53, 69.5, 55, 71, 63, 76.5, 65.5, 69, 75, 76, 57, 70.5,
+ 71.5, 56, 81.5, 69, 59, 67.5, 61, 68, 59.5, 56.5, 73,
+ 61, 72.5, 71.5, 59.5, 74.5, 63)

h1<- graph.freq(weight,col="yellow",frequency=1,las=2,xlab="h1")

print(summary(h1),row.names=FALSE)
```


##3. Average
```{r}
df.saea.clean%>%
          filter(terminado=="si"&Estado=="En Proceso"&grepl("INDRA",Especialista.de.Estado)&!is.na(serv_TI))%>% 
          group_by(Linea.de.Servicio=serv_TI,Fecha.final)%>%
          summarise(horas=round((sum(Tiempo)/60),2))%>%
          group_by(Linea.de.Servicio) %>%
          summarise(Mean = mean(horas))
```



