


```{r}
if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}

```

Grafic
```{r install}

if (!require("visdat")) {
  install.packages("visdat", repos="http://cran.rstudio.com/") 
  library(visdat)
}

if (!require("naniar")) {
  install.packages("naniar", repos="http://cran.rstudio.com/") 
  library(naniar)
}

if (!require("rpart")) {
  install.packages("rpart", repos="http://cran.rstudio.com/") 
  library(rpart)
}


if (!require("DataExplorer")) {
  install.packages("DataExplorer", repos="http://cran.rstudio.com/") 
  library(DataExplorer)
}

if (!require("rpart.plot")) {
  install.packages("rpart.plot", repos="http://cran.rstudio.com/") 
  library(rpart.plot)
}

```

Dataset
```{r}
if (!require("nycflights13")) {
  install.packages("nycflights13", repos="http://cran.rstudio.com/") 
  library(nycflights13)
}
```

load data
```{r}
#From 
df.dqa<-read.csv("C:/R-Source/hot-dog-contest-winners.csv")

#from library
data_list <- list(airlines, airports, flights, planes, weather)
```

Now let's merge all tables together for a more robust dataset for later sections.
```{r}
merge_airlines <- merge(flights, airlines, by = "carrier", all.x = TRUE)
merge_planes <- merge(merge_airlines, planes, by = "tailnum", all.x = TRUE, suffixes = c("_flights", "_planes"))
merge_airports_origin <- merge(merge_planes, airports, by.x = "origin", by.y = "faa", all.x = TRUE, suffixes = c("_carrier", "_origin"))
final_data <- merge(merge_airports_origin, airports, by.x = "dest", by.y = "faa", all.x = TRUE, suffixes = c("_origin", "_dest"))
```


sURCE: 
*An Introduction to Data Profiling, Felix Naumann
*https://cran.r-project.org/web/packages/DataExplorer/vignettes/dataexplorer-intro.html
*https://www.rdocumentation.org/packages/DataExplorer/versions/0.6.0

I.SINGLE COLUM
==============================


#1.CARDINALITY

##1.1.Number of values by table

##1.2.Number of distinct values

Charging colum names to a "list"
```{r}
#de.1 is a data frame
lista1<-names(final_data)


class(lista1)
for (i in 1:nrow(final_data)){
y<- NROW(unique(final_data[,i]))
}

iris.num <- iris[, -5]
apply(iris.num, 2, min)
```



```{r}
for (i in 1:nrow(final_data)){
#j <- i + 10
print(paste(col=colnames(final_data)[i],NROW(unique(final_data[,i])),sep = ":"))
}
```



##1.3.Number of NULLs (Missing values)
###1.General
1.1.Usign dataExplorer
```{r}
plot_missing(final_data)
```


1.2.Using "visdat" library)
```{r all}
#vis_miss: The function vis_miss provides a summary of whether the data is missing or not. It also provides the amount of missings in each columns.
vis_miss(airquality)
```

###2.Look into the missing value columns
```{r}
#geom_point
ggplot(airquality, 
       aes(x = Solar.R, 
           y = Ozone)) + geom_miss_point()
#faceting by no-missing column
ggplot(airquality, 
       aes(x = Solar.R, 
           y = Ozone)) + 
  geom_miss_point()+ 
  facet_wrap(~Month)
```

###3.MIssingness distribution 
3.1.Curve(SHADOWS)
```{r}
#shadow(): functions provide a way to keep track of missing values. The as_shadow function creates a dataframe with the same set of columns, but with the column names added a suffix _NA
as_shadow(airquality)
aq_shadow <- bind_shadow(airquality)
glimpse(aq_shadow)

ggplot(aq_shadow,
       aes(x = Temp,
           colour = Ozone_NA)) + 
  geom_density()
```

3.2.missingness tree dessition
```{r}
airquality %>%
  add_prop_miss() %>%
  head()

library(rpart)
library(rpart.plot)
airquality %>%
  add_prop_miss() %>%
  rpart(prop_miss_all ~ ., data = .) %>%
  prp(type = 4, extra = 101, prefix = "Prop. Miss = ")
```




##MIN and MAX value


#2.UNIQUENESS AND KEYS
##2.1. Unique column
 Only unique values (scales library)
```{r}
###3.Identify unicicity 
num.total<-nrow(serv.saea)

dem.eventos<-n_distinct(serv.saea$`Demanda del Servicio Eventos`, na.rm = FALSE)/num.total
#using  "percent()" from "scales" package, to plot ot will be disable
print(percent(dem.eventos))
```
 
 
##2.2.Unique column combination
??? Only unique value combinations
??? Minimality: No subset is unique


#2.3.Key candidate
??? No null values
??? Uniqueness and non-null in one instance does not imply key: Only human can specify keys (and foreign keys)

#2.4.Meaning of NULL values for each column



#3.PATTERN & DATATYPE

##3.1.Datatype by table
Tree: HOw many record of String vs. number vs. date exit?(Using "DataExplorer" library)
```{r}
#library(DataExplorer)
plot_str(data_list)
```
Table (Using "visdat" library)
```{r}
#vis_dat: Amount, class, missing
        #visualises the whole dataframe at once, and provides information about the "class"" of the data input into R, as well as whether the data is "missing"" or not
vis_dat(airquality)
```



##Categorical vs. continuous
Resume of discrete and continuous column ("DataExplorer" library)
```{r}
introduce(final_data)
```

Split data into discrete and continuous parts ("DataExplorer" library)
```{r}
output <- split_columns(iris)
output$discrete
output$continuous
output$num_discrete
output$num_continuous
output$num_all_missing
```



##SQL data types: () CHAR, INT, DECIMAL, TIMESTAMP, BIT, CLOB,etc)
##Domains VARCHAR(12) vs. VARCHAR (13)
## XML data types() More fine grained)
## Regular expressions (\d{3})-(\d{3})-(\d{4})-(\d+)
## Semantic domains (Adress, phone, email, first name)

#4.Value Distributions

##4.1.Bar chart
```{r}
plot_bar(final_data)
```


##4.2.Histogram
```{r}
colors = c("red","green3","blue")
numcols <- length(names(df.dqa))
par(mfrow=c(3,5))
for(i in 1:numcols){
if(is.factor(df.dqa[,i])){
if( as.integer(nlevels(df.dqa[,i]) <= 20) )
plot(df.dqa[,i],main=names(df.dqa)[i],col=colors)
}
else{hist(df.dqa[,i],main=names(df.dqa)[i],xlim=c(0,300000),breaks=100,xlab=names(df.dqa)[i],col=colors)
}
}
```


```{r basis}
# 1. HISTOGRAM FOR PARAMETRIC
hist(airquality$Solar.R, breaks=10, col="red")


# 2. HISTOGRAM FOR NO PARAMETRIC
ggplot(data=airquality, aes(airquality$Solar.R)) + 
  geom_histogram(breaks=seq(20, 300, by =20), 
                 col="red", 
                 aes(fill=..count..))


```




Visualising missings in variables
```{r}
#gg_miss_var(): shows the number of missing values in each variable in a dataset
gg_miss_var(airquality)+ labs(y = "Look at all the missing ones")

#gg_miss_case(): This plot shows the number of missing values in each case. It is powered by the miss_case_summary() function.
gg_miss_case(airquality) + labs(x = "Number of Cases")

#gg_miss_fct():shows the number of missings in each column, broken down by a categorical variable from the dataset.
View(riskfactors)
gg_miss_fct(x = riskfactors, fct = marital)

#
# using group_by
library(dplyr)
riskfactors %>%
  group_by(marital) %>%
  miss_var_summary()

#gg_miss_span(): This plot shows the number of missings in a given span, or breaksize, for a single selected variable. In this case we look at the span of hourly_counts from the pedestrian dataset.
miss_var_span(pedestrian, hourly_counts, span_every = 3000)
gg_miss_span(pedestrian, hourly_counts, span_every = 3000)+labs(x = "custom") + theme_light()


```









USING VIM PACKAGE
```{r}
install.packages("VIM")
library(VIM)
data(tao)
summary(tao)
View(tao)

#matrixplot() function will display the entire dataset as a set of rectangles to represent each cell. Similar values in the data will have similar grayscale values, and NAs will show up clearly in red:
matrixplot(tao)

matrixplot(airquality)

#To complement the detailed picture given by matrixplot ,you can get a numerican d visual summary of missing values with the aggr function:
tao_aggr = aggr(tao)
tao_aggr

#You can compare missingness across fields with histograms, such as the distribution of missing humidity values by corresponding air temperature:
histMiss(tao[5:6])

#You can flip the variables to see the opposite distribution:
histMiss(tao[c(6,5)])


#You can show the same data in a scatterplot for more details:
marginplot(tao[5:6])


#This can be expanded to visualizing pairwise comparisons with the marginmatrix function:
marginmatrix(tao[4:6])

```


II.MULTIPLE COLUM
==============================
#1.Uniqueness and keys



#2.Inclusion and foreign key dependency



#3.Functional dependencies


#4.Conditional and approximate dependencies
