---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---


```{r install}

if (!require("visdat")) {
  install.packages("visdat", repos="http://cran.rstudio.com/") 
  library(visdat)
}

if (!require("naniar")) {
  install.packages("naniar", repos="http://cran.rstudio.com/") 
  library(naniar)
}

if (!require("rpart")) {
  install.packages("rpart", repos="http://cran.rstudio.com/") 
  library(rpart)
}


if (!require("DataExplorer")) {
  install.packages("DataExplorer", repos="http://cran.rstudio.com/") 
  library(DataExplorer)
}

if (!require("rpart.plot")) {
  install.packages("rpart.plot", repos="http://cran.rstudio.com/") 
  library(rpart.plot)
}


if (!require("ggplot2")) {
  install.packages("ggplot2", repos="http://cran.rstudio.com/") 
  library(ggplot2)
}

if (!require("data.table")) {
  install.packages("data.table", repos="http://cran.rstudio.com/") 
  library(data.table)
}

if (!require("rmarkdown")) {
  install.packages("rmarkdown", repos="http://cran.rstudio.com/") 
  library(rmarkdown)
}
```


```{r}
config <- list(
"introduce" = list(),
"plot_str" = list(
"type" = "diagonal",
"fontSize" = 35,
"width" = 1000,
"margin" = list("left" = 350, "right" = 250)
),
"plot_missing" = list(),
"plot_histogram" = list(),
"plot_bar" = list(),
"plot_correlation" = list("use" = "pairwise.complete.obs"),
"plot_prcomp" = list(),
"plot_boxplot" = list(),
"plot_scatterplot" = list()
)
# Create report
create_report(iris)
create_report(airquality, y = "Ozone")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
