---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---



```{r install}

if (!require("visdat")) {
  install.packages("visdat", repos="http://cran.rstudio.com/") 
  library(visdat)
}

if (!require("naniar")) {
  install.packages("naniar", repos="http://cran.rstudio.com/") 
  library(naniar)
}

if (!require("rpart")) {
  install.packages("rpart", repos="http://cran.rstudio.com/") 
  library(rpart)
}


if (!require("DataExplorer")) {
  install.packages("DataExplorer", repos="http://cran.rstudio.com/") 
  library(DataExplorer)
}

if (!require("rpart.plot")) {
  install.packages("rpart.plot", repos="http://cran.rstudio.com/") 
  library(rpart.plot)
}


if (!require("ggplot2")) {
  install.packages("ggplot2", repos="http://cran.rstudio.com/") 
  library(ggplot2)
}

if (!require("data.table")) {
  install.packages("data.table", repos="http://cran.rstudio.com/") 
  library(data.table)
}

if (!require("rmarkdown")) {
  install.packages("rmarkdown", repos="http://cran.rstudio.com/") 
  library(rmarkdown)
}
```


s
```{r}

data("diamonds", package = "ggplot2")
# Set some missing values
diamonds2 <- data.table(diamonds)
for (j in 5:ncol(diamonds2)) {
set(diamonds2,
i = sample.int(nrow(diamonds2), sample.int(nrow(diamonds2), 1)),j,value = NA_integer_)
}
# Create customized report for diamonds2 dataset
create_report(
data = diamonds2,
output_file = "report.html",
output_dir = getwd(),
y = "price",
config = list(
"introduce" = list(),
"plot_missing" = list(),
"plot_histogram" = list(),
"plot_density" = list(),
"plot_bar" = list("with" = "carat"),
"plot_correlation" = list("use" = "pairwise.complete.obs"),
"plot_prcomp" = list(),
"plot_boxplot" = list("by" = "carat"),
"plot_scatterplot" = list("by" = "carat")
),
html_document(toc = TRUE, toc_depth = 6, theme = "flatly")
)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
