---
title: "R Notebook"
output: html_notebook
---

```{r}
if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}

if (!require("tidyr")) {
  install.packages("tidyr", repos="http://cran.rstudio.com/") 
  library(tidyr)
}

if (!require("stringr")) {
  install.packages("stringr", repos="http://cran.rstudio.com/") 
  library(stringr)
}

```

```{r}
#SOurce:R for Data Science - Hadley Wickham; Data Wrangling with R
#Determine which strings match a pattern.
#Find the positions of matches.
#Extract the content of matches.
#Replace matches with new values.
#Split a string based on a match.
```


I.STRING MANIPULATION (inside a word)
====================

#1.UPPER, lOWER
###1 ways to turn all into "upper" ("casefold" or "toupper")
```{r}
#casefold
casefold(iris$Species,upper = TRUE)

#toupper
toupper(iris$Species)
```


##2. Abreviate
```{r}
#We use "Abbreviate" function
abbreviate(iris$Species,4,method = "left.kept")

```

###3.Lenght of string
```{r}
str_length("abcdef")
str_length(iris$Species)
```



2.MATCHES(stringr,stringr PACKAGE)
===================
##1 matching a word inside a column by using DPLYR & STRINGR
```{r}
ans2.total%>%filter(str_detect(ans2.total$descrip_def,"UAT"))
```


##2. Detect  string giving the values (str_view from "stringr" library)
```{r}
x1 <- c("apple", "banana", "pear", "arandano")
str_view(x1,"^a")
str_view(x1,"a$")
str_view(x1,".a.")

x2 <- c("apple pie", "apple", "apple cake")
str_view(x2,"apple")
str_view(x2,"apple$")

```


##3.Detect  string giving True-false of the matching
###A. str_detect from "stringr" PACKAGE)
```{r}
  #Boolean conditin of detecting "e" inside each word
  str_detect(x1,"app")
  # How many common words start with t?
  sum(str_detect(x1,"e"))

#6.Find Matches with "PATTERNS" uppercase or lowercase forms
    #Now we are going to use a "fruit" from "stringr" library
  stringr::fruit
  
  str_view(fruit,"nana")
  #usign regex let us use ignore_case = TRUE, to match either their uppercase or lowercase forms
  str_view(fruit,regex("nana",ignore_case = TRUE))
```



###C. grepl from "base"
```{r}
mtcars$type <- rownames(mtcars)
filter(mtcars, grepl('Toyota|Mazda', type))

#the complement
dplyr::filter(mtcars, !grepl('Toyota|Mazda', type))
```


###D.PAste
```{r}

paste(mtcars$mpg, ".csv", sep="")
```

##2.3.REPLACING SUBTRINGS BY EXACT VALUES(stringr PACKAGE)
```{r}
#1.First just replacing the first vowel fron each word
  str_replace(x1,"[aeiou]","-")
  #Now replacing the all vouwels fron each word
  str_replace_all(x1,"[aeiou]","-")
  #With str_replace_all() you can perform multiple replacements by supplying a named vector

#2.Now replace and assign
  #asing to a object called "vector1"
  vector1<-head(iris,150)
  vector1$Species<-str_replace_all(vector1$Species,"to","MA")

  #If a want to replace more than one value
  vector1$col1<-str_replace_all(vector1$Species,c("se"="one","MA"="two","sa"="three"))
  
    
#3. remove accents
  data1$newcol<-chartr("?????", "aeiou",data1$Proceso_afectado) 
```


##2.4. EXTRACT CHARACTERS BY POSITION-SUBTRINGS
###1.From a begin to end position
```{r}
#Subtring the colum
substr("abcdef", 2, 4)
```
###2. From a begin to infinite
```{r}
substring("abcdef", 2)
```



