---
title: "R Notebook"
output: html_notebook
---

```{r}
if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/") 
  library(dplyr)
}


if (!require("stringr")) {
  install.packages("stringr", repos="http://cran.rstudio.com/") 
  library(stringr)
}

```

```{r}
#SOurce:R for Data Science - Hadley Wickham; Data Wrangling with R


#Determine which strings match a pattern.
#Find the positions of matches.
#Extract the content of matches.
#Replace matches with new values.
#Split a string based on a match.
```


I.STRING MANIPULATION (inside a word)
====================
#1.REPETITION AND SEQUENCES
##.Create vectors of strings
```{r}
## a) ( 1,2,3, ??? ,19,20) 
v1<-c(1:20)

## b) ( 20,19, ??? ,2,1 )
v2<-c(20:1)

## c) ( 1,2,3, ??? ,19,20,19,18, ??? ,2,1 )
v3<-c(1:20,19:1)

## d) (4,6,3) y asignarlo a la variable tmp
tpm<-c(4,6,3)

## e) ( 4,6,3,4,6,3, ??? ,4,6,3 ) donde el numero 4 aparece 10 veces.
v4<-rep.int(tpm,times=10)

#f) ( 4,6,3,4,6,3, ??? ,4,6,3,4 ) donde hay 11 ocurrencias del n?mero 4, 10 ocurrencias del n?mero 6 y 10 ocurrencias del n?mero 3
v5<-c(v4,4)
##g) ( 4,4, ??? ,4,6,6, ??? ,6,3,3, ??? ,3 ) donde hay 10 ocurrencias del numero 4, 20 ocurrencias del numero 6 y 30 ocurrencias del numero 3
v6<-rep(4,times=10)
v7<-rep(6,times=20)
v8<-rep(3,times=30)
v9<-c(v6,v7,v8)
```

##Create vectors of date
```{r}
#1.Create a dataframe of dates 
df.dates<-seq(as.Date("1910/1/1"), as.Date("1999/1/1"), "years")
```

##3. Add a column with sequential values
```{r}
df.iris<-iris
df.iris$id <- seq.int(nrow(iris))

```


#2.UPPER, lOWER
### 2 ways to turn all into "upper" ("casefold" or "toupper")
```{r}
#casefold
casefold(iris$Species,upper = TRUE)


#toupper
toupper(iris$Species)
```


##3. Abreviate
  We use "Abbreviate" function
```{r}
abbreviate(iris$Species,4,method = "left.kept")

```

#4.DETEC MATCHES(stringr,stringr PACKAGE)
##4.1.Detect  string giving the values (str_view from "stringr" library)
```{r}
x1 <- c("apple", "banana", "pear", "arandano")
str_view(x1,"^a")
str_view(x1,"a$")
str_view(x1,".a.")

x2 <- c("apple pie", "apple", "apple cake")
str_view(x2,"apple")
str_view(x2,"apple$")

```


##4.2.Detect  string giving True-false of the matching
### str_detect from "stringr" PACKAGE)
```{r}
  #Boolean conditin of detecting "e" inside each word
  str_detect(x1,"app")
  # How many common words start with t?
  sum(str_detect(x1,"e"))

#6.Find Matches with "PATTERNS" uppercase or lowercase forms
    #Now we are going to use a "fruit" from "stringr" library
  stringr::fruit
  
  str_view(fruit,"nana")
  #usign regex let us use ignore_case = TRUE, to match either their uppercase or lowercase forms
  str_view(fruit,regex("nana",ignore_case = TRUE))
```

### Similar a SQL "like" function "stringr" and "dplyr"
```{r}
fruta$dato=as.data.frame(fruit)

fruta %>%
  filter(str_detect(dato, 'black|blu'))
```


### grepl from "base"
```{r}
mtcars$type <- rownames(mtcars)
filter(mtcars, grepl('Toyota|Mazda', type))

#the complement
dplyr::filter(mtcars, !grepl('Toyota|Mazda', type))
```





#PAste
```{r}

paste(mtcars$mpg, ".csv", sep="")
```


```{r}
file.path(directory, paste(id, ".csv", sep=""))
```



#4.3.REPLACING SUBTRINGS BY EXACT VALUES(stringr PACKAGE)
```{r}
#1.First just replacing the first vowel fron each word
  str_replace(x1,"[aeiou]","-")
  #Now replacing the all vouwels fron each word
  str_replace_all(x1,"[aeiou]","-")
  #With str_replace_all() you can perform multiple replacements by supplying a named vector

#2.Now replace and assign
  #asing to a object called "vector1"
  vector1<-head(iris,150)
  vector1$Species<-str_replace_all(vector1$Species,"to","MA")

  #If a want to replace more than one value
  vector1$col1<-str_replace_all(vector1$Species,c("se"="one","MA"="two","sa"="three"))
  
    
#3. remove accents
  data1$newcol<-chartr("?????", "aeiou",data1$Proceso_afectado) 
```


#4.4. EXTRACT CHARACTERS BY POSITION-SUBTRINGS
##1.From a begin to end position
```{r}
#Subtring the colum
substr("abcdef", 2, 4)
```
##2. From a begin to infinite
```{r}
substring("abcdef", 2)
```



```{r}
str_length
```



SEPARATE COLUMS
=========================
3.1.Separate columns in R by space delimiter
```{r}
#Creating DATA FRAME
df<-data.frame(col1=c("SOA ABC BD","MND OOL POL"),col2=c(2,8))
#Removing Whitespace using: \\s+
demo<-df%>%
    separate(col1,c("comp1","comp2","comp3"),sep = "\\s+",remove = FALSE)
```


3.2.
```{r}
#1.First import  
ss<-read.csv("C:/R-Source/ss.csv")

#Split with special character(*,/,(, etc)), then use "separate"
  ara.1 <- ara.df%>%  
  separate(Jerarquia,c("Macroservicio","Servicio_ti","Subservicio","Categoria1","Categoria2","Categoria3","Categoria4"), sep ='\\.|\\.|\\.|\\.|\\.|\\.', remove = FALSE)
  
  

```

DPLYR & STRINGR
================
# matching a word inside a column
```{r}
ans2.total%>%filter(str_detect(ans2.total$descrip_def,"UAT"))
```

```{r}
#. You can find the complete list here, but here we can use '[:alpha:]' to represent all alphabet characters including both upper and lower cases.
flight %>%
  select(FL_DATE, CARRIER, ORIGIN, ORIGIN_CITY_NAME, ORIGIN_STATE_ABR, DEP_DELAY, DEP_TIME, ARR_DELAY, ARR_TIME) %>%
filter(str_detect(ORIGIN_CITY_NAME, "New[:alpha:]+rk")) %>%
count(ORIGIN_CITY_NAME)
```

