---
title: "R Notebook"
output: html_notebook
---

```{r}
#https://data-se.netlify.com/2017/02/13/data_cleansing/
#https://dabblingwithdata.wordpress.com/2017/12/20/my-favourite-r-package-for-frequency-tables/
#https://www.youtube.com/watch?v=qGqAhlFzkSU
#https://www.youtube.com/watch?v=KbLuSEINxdM

```



```{r}
install.packages("janitor")
install.packages("epiDisplay")
install.packages("GGally")
install.packages("summarytools")
```

```{r}
library(janitor)
library(dplyr)
```



get_dupes: Get rows of a data.frame with identical values for the specified vari-
ables
```{r}
get_dupes(mtcars, mpg, hp)
# or called with the magrittr pipe %>% :
mtcars %>% get_dupes(wt)
```

tabyl
==============
can be called two ways:

1.On a vector
 when tabulating a single variable - e.g., tabyl(roster$subject)
```{r}
mtcars%>%
    tabyl(cyl)
```

2.On a data.frame
    specifying 1, 2, or 3 variable names to tabulate : roster %>% tabyl(subject, employee_status).Here the data.frame is passed in with the %>% pipe; this allows tabyl to be used in an analysis pipeline
2.1.Two variables:    
```{r}
roster %>%
  filter(hire_date > as.Date("1950-01-01")) %>%
  tabyl(employee_status, full_time)
```
    



```{r}
library(lubridate)
df<-data.frame(x1=c("2017-03-23","2017-03-24"),x2=c("2017-03-24","2017-03-28"))
df[]<-lapply(df,ymd)
df$x2-df$x1
```




crosstab
===================
displaying either frequencies or percentages calculated by row, column, or overall.
```{r}
mtcars %>%
  filter(cyl == 6) %>%
  crosstab(vs, carb)

crosstab(mtcars$mpg,mtcars$cyl)
```




```{r}
#https://dabblingwithdata.wordpress.com/2018/02/26/r-packages-for-summarising-data-part-2/
library(epiDisplay)
tab1(mtcars$cyl, sort.group = "decreasing", cum.percent = TRUE)
```
    


s
=================

```{r}
library(GGally)
ggpairs(mtcars)
```



```{r}
library(summarytools)
freq(iris$Species, style = "rmarkdown", order = "freq")
```

Identify variables with unique values
==================



a
====================



```{r}
mtcars %>%
tabyl(am, cyl) %>%
adorn_percentages("col") %>%
adorn_pct_formatting() %>% #Format a data.frame of decimals as percentages.
adorn_ns(position = "front") #Add underlying Ns to a tabyl displaying percentages
```


```{r}
mtcars %>%
tabyl(am, cyl) %>%
adorn_percentages("col") %>%
adorn_pct_formatting()
```


