---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
book "r for data sceince""

CHAPTERS


Install Libraries
```{r Install Libraries}
install.packages("tidyverse")
install.packages("tidyr")
install.packages("dplyr")
install.packages("grid")
install.packages("tibble")
install.packages("stringr")
install.packages("sqldf")

#it's better
install.packages(c("nycflights13", "Lahman"))
install.packages("lubridate")
```

Calling Libraries
```{r Calling Libraries}
library(tidyverse)
library(dplyr)
library(grid)
library(tibble)
library(tidyr)
library(stringr)
library(readr)
library(sqldf)
library(ggplot2)
library (xlsx)
library(rJava)
library(xlsxjars)

#for working with dates
library(lubridate)


```

I.IMPORTING
```{r}
#We can read from all "sheet" of any excel file
data1<-read.xlsx ("C:/R-Source/procesos1.xlsx", sheetName = "Hoja1")
View(data1)
```

II.TIDY
baSIC
```{r}
#1.CREATE COLUM




#using loops with dataframes

> myDf <- head(iris)
> myDf
Sepal.Length Sepal.Width Petal.Length Petal.Width Species
1          5.1         3.5          1.4         0.2  setosa
2          4.9         3.0          1.4         0.2  setosa
3          4.7         3.2          1.3         0.2  setosa
4          4.6         3.1          1.5         0.2  setosa
5          5.0         3.6          1.4         0.2  setosa
6          5.4         3.9          1.7         0.4  setosa
> nRowsDf <- nrow(myDf)
> for(i in 1:nRowsDf){
+ print(myDf[i,4])
+ }
[1] 0.2
[1] 0.2
[1] 0.2
[1] 0.2
[1] 0.2
[1] 0.4




#2.Delete Colum

```

06.Abbreviate names for "Macroservicio" and "Servicio_ti"
  We use "Abbreviate" function
```{r}
dem$Macroservicio<-abbreviate(dem$Macroservicio,20,method = "left.kept")
View(dem)

```

tibble
```{r}
eje<-tibble(x=1:6,y=2,z=x^3+y)
eje$x
eje[["x"]]
eje[[1]]

#ANCHOR
#Using "srt_view" from "htmlwidgets" library
x1 <- c("apple", "banana", "pear", "arandano")
str_view(x1,"^a")
str_view(x1,"a$")
str_view(x1,".a.")

x2 <- c("apple pie", "apple", "apple cake")
str_view(x2,"apple")
str_view(x2,"apple$")

#2.REPETITION
```

3.DETEC MATCHES
```{r}
  #Boolean conditin of detecting "e" inside each word
  str_detect(x1,"app")
  # How many common words start with t?
  sum(str_detect(x1,"e"))

#6.Find Matches with "PATTERNS" uppercase or lowercase forms
    #Now we are going to use a "fruit" from "stringr" library
  stringr::fruit
  
  str_view(fruit,"nana")
  #usign regex let us use ignore_case = TRUE, to match either their uppercase or lowercase forms
  str_view(fruit,regex("nana",ignore_case = TRUE))
```



III.OPERATIONS (WITHOUT TRANSFORMING DATA)
=====================
Charging colum names to a "list"
```{r}
#de.1 is a data frame
lista1<-names(de.1)
#or viceversa also run, as for pareto bean
```

tOTAL SUM
```{r}
#First charging a bean 
  be.paret.app<-de.1 %>%
    group_by(criterio=sistema_relacionado) %>%
    summarize(cantidad=n())%>%
    filter(!is.na(criterio))%>%
    #THIS IS for figure out the total as a new colum
    mutate(s=sum(cantidad))
```
order rows using "ARRANGE"
```{r}

  be.bar.ser.app<-de.1 %>%
    group_by(periodo=sistema_relacionado,criterio=Sigla_Macroservicio) %>%
    summarize(cantidad=n())%>%
    filter(!is.na(criterio)&periodo%in%c("NSP","NSTD","APJ","SIGA","NSBR","MCCIA","OYR","APPENS","REFLEX","SAA"))%>%
    arrange(desc(cantidad))
```



III.TRANSFORM
=========================
Libraries: tibble,stringr
1.Basic update
```{r}
ara.dqa.dep$area<-ara.dqa.dep$Unid_org
#Remove just one "vocal"
ara.dqa.dep$area<-str_replace_all(ara.dqa.dep$area,"?","o")
# remove accents
ara.dqa.dep$area<-chartr("?????", "aeiou",ara.dqa.dep$area)   

#You use ifelse when you have to replace all 
ara.dqa.dep$area<- ifelse(str_detect(ara.dqa.dep$area,regex("Direccion de Inversiones")),"DIN",ara.dqa.dep$area)
#if you just want to update same pattern without affect the others one, do that
ara.dqa.dep$area<- ifelse(str_detect(ara.dqa.dep$area,regex("Direccion de Inversiones")),"DIN",ara.dqa.dep$area)
```


2.Advance update, using comparison


update colums
```{r  update colums}
#we cannot use a list of patterns whe we use "grepl"
d$Flat <- ifelse(grepl("Apt",d$Flat),"Apartment","House")

rtc.1$macroservicio[rtc.1$Tipo=="InformeTecnico"|rtc.1$Tipo_de_Requerimiento=="Informes tecnicos"]<-"MS11-INFORMES TECNICOS"
rtc.1$macroservicio[rtc.1$Tipo%in%c("InformeTecnico","Informes tecnicos")]<-"MS11-INFORMES TECNICOS"

#This is the best wa  y
rtc.1<-rtc.1 %>%
  mutate(fecha_de_cierre= ifelse(grepl("Cerrado",Estado),Fecha_de_modificacion,NA)) 


```

update one table from another, prior to Compare 1 columns
```{r}
#1.Populate from "CAtegoria 2",prior to compare against Anexo5_sistemas_Aranda
ara.1$newSis_Cat2<-ifelse(match(ara.1$Categoria2,ara.sis$Sigla_sist),ara.1$Categoria2,ara.1$sistema_relacionado)
rtc.1$proceso_relacionado<-rtc.proc.nonull$Proceso_OIP[match(rtc.1$Proceso_afectado,rtc.proc.nonull$Proceso_afectado)]
```


update one table from another, prior to Compare 2 columns
```{r}
rtc.1$estado_req<-ifelse(match(rtc.1$macroservicio,rtc.est.gen$Macroservicio_aux)&match(rtc.1$Estado,rtc.est.gen$Estado_req_aux),rtc.est.gen$Estado_generico,NA)
```


Upper 
2 ways to turn all into "upper" ("casefold" or "toupper")
```{r}
#casefold
ara.dqa.dep$SubArea<- casefold(ara.dqa.dep$SubArea,upper = TRUE)

#toupper
ara.dqa.dep$Area<-toupper(ara.dqa.dep$Area)

```



```{r}
#
ara.df.2<-ara.df.1 %>%
          mutate(sis_cate= ifelse(grepl("NSP",ara.df.1$Categoria3),"NSP","NA")) 

# finding "SISTEMAS" in the colum "Categoria3". 
  ara.df.1 %>%
  filter(ara.df.1$Categoria3 %in% c("NSBR","NSP"))%>%
  select(Servicio_ti,Categoria3)
```


```{r}
#4.REPLACING MATCHES
  #First just replacing the first vouwel fron each word
  str_replace(x1,"[aeiou]","-")
  #Now replacing the all vouwels fron each word
  str_replace_all(x1,"[aeiou]","-")
  #With str_replace_all() you can perform multiple replacements by supplying a named vector
  x3 <- c("1 house", "2 cars", "3 people")
  str_replace_all(x3,c("1"="one","2"="two","3"="three"))
  
  #BUT FOR DATAFRAME IS DIFFERENT
  ara.df.2$Servicio_ti<-str_replace_all(ara.df.2$Servicio_ti,"S025","S25")
  
  
  #Now we are going to use a "sentences" from "stringr" library
  stringr::sentences
  length(sentences)
  head(sentences)
  #I flip the order of the second and third words
  sentences%>%
          str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2")%>%
          head(5)
#5.Split and separete
  #Split with special character(*,/,(, etc)), then use "separate"
  ara.1 <- ara.df%>%  
  separate(Jerarquia,c("Macroservicio","Servicio_ti","Subservicio","Categoria1","Categoria2","Categoria3","Categoria4"), sep ='\\.|\\.|\\.|\\.|\\.|\\.', remove = FALSE)
  #SPlit with characters ("and","te", etc), then use "split"
  str_split(fruits, " and ", simplify = TRUE)
  
  
```


.  "date"
```{r "date"}
###a.ARANDA: Creating new field "fecha" for Analysis  with the form "A?O-MES"
ara.1$fecha<-format(as.Date(ara.1$fecha_de_solicitud,format="%Y-%m-%d"),"%Y-%m")
View(ara.1)

rtc_tem<-rtc.1%>%
  select(fecha_de_solicitud,fecha_de_cierre,Servicio_ti)
View(rtc_tem)

###b.RTC: Now Turn "fecha_de_solicitud" into date using "LUBRIDATE" library
rtc_tem$fecha_de_solicitud<-mdy_hm(rtc_tem$fecha_de_solicitud)
View(rtc_tem)

mdy_hm("01/31/2017 08:01")

```


Operations 
1.Sum
```{r}
rtc.1$esfuerzo_total<-as.numeric(rtc.1$Esfuerzo_SFS)+as.numeric(rtc.1$Esfuerzo_GIC)
```

VII.SAVE
Libraries: write,xlsx (it comes with "xlsxjars" and "rJava" libraries)
```{r VII.SAVE}
#use write.xlsx from "WriteXLS" library
#Warning: lESS THAN 66000 ROWS
WriteXLS("ara.2","C:/R-Source/back_rtc_1.xls")


#we'll use write.xlsx from "xlsx" library
write.xlsx(rtc.proc,file ="C:/R-Source/Anexo1_EquivProcesosNoNull_RTC.xlsx")
```
